---
title: 'STAT 500: HW1'
author: "Jasmine Mou"
date: "9/18/2017"
output:
  pdf_document: default
---

1. Classify the following variables as binary, discrete, or continuous. Also classify them as qualitative or quantitative. Example: Age in years. Answer: Continuous and quantitative
(a) Time in terms of AM or PM.  
*binary, as the time should either be in AM or in PM, and quanlitative. *

(b) Angles as measured in degrees between 0 and 360.  
*continuous and quantitative.*

(c) Bronze, Silver, and Gold medals as awarded at the Olympics.  
*discrete and qualitative.*

(d) Height above sea level.  
*continous and quantitative.* 

(e) Number of patients in a hospital.  
*discrete and quantitative.*


2. Do exploratory data analysis for ”Pima Indian” data.
```{r init}
library(faraway)
data(pima)
```

(a) Find the dimension of the data (p and n).
```{r (a)}
dim <- dim(pima)
n <- dim[1]
p <- dim[2]
```
*Thus the dimension of data is n = `r n` and p = `r p`.*  

(b) Find the numerical summaries of each variables (min, 1Q, median, mean, 3Q, max).  
```{r (b), warning=FALSE}
t_summary <- apply(pima, 2, function(x) summary(x))
library(knitr)
kable(t_summary)
```

(c) Change some of the 0s to **NA** and find the numerical summaries of each variables again.   
*Here we change `glucose`, `triceps`, and `insulin`'s 0s to NA.*  
```{r (c)}
pima$glucose[pima$glucose==0] = NA
pima$triceps[pima$triceps==0] = NA
pima$insulin[pima$insulin==0] = NA
```
*In the new summary, we notice the update of `Min.` values for these 3 variables.*  
```{r (c) cont}
t_summary <- apply(pima, 2, function(x) summary(na.omit(x)))
kable(t_summary)
```

(d) Which observation has the largest ”diastolic blood pressure”? Give the row number.  
```{r (d)}
index_max_diastolic <- pima$diastolic==max(pima$diastolic, na.rm=TRUE)
pima[index_max_diastolic,]
```
*The row number is `r which(index_max_diastolic)`.*   

(e) Draw the graphical summaries (histogram and boxplot) for a single variable ”diastolic blood pressure”. **(hint: you may want to use *par(mfrow = c(1,2))* )**  
```{r (e)}
par(mfrow=c(1,2))
hist(pima$diastolic, xlab="diastolic", ylab="Frequency", main="Histogram of diastolic")
boxplot(pima$diastolic, ylab="diastolic", main="Boxplot of diastolic")
```

(f) Draw the graphical summaries (histogram and boxplot) for two variables; ”test” and ”bmi”. You must present at least one graph showing the relationship between two variables. **(hint: you may want to use *par(mfrow = c(1,3)) and factor()* )**  
*When viewing `test` as continuous variables (0,1), we can summarize it with histogram. However, the clustered representation of `test` and `bmi` is not clear enough.*   
```{r (f)}
par(mfrow=c(1,3))
hist(pima$test, xlab="test", ylab="Frequency", main="Histogram of test (cont.)")
boxplot(pima$bmi, xlab="BMI", ylab="Frequency", main="Boxplot of bmi (cont.)")
plot(x=pima$test, y=pima$bmi, xlab="test", ylab="BMI", main="test vs bmi")
```

*Thus we factorize the `test` first before summarize the relationship between `test` and `bmi`. When `test` is converted to categorical variable rather than the original continous variable, histogram doesn't apply in representation of single variable `test`, and we use barplot instead.* **[reference link](https://stackoverflow.com/a/24600821)**
```{r (f) cont}
par(mfrow=c(1,3))
pima$test=factor(pima$test, labels=c("negative", "positive"))
barplot(table(pima$test), main="Barplot of test (cat.)")
boxplot(pima$bmi, xlab="BMI", ylab="Frequency", main="Boxplot of bmi (cont.)")
plot(x=pima$test, y=pima$bmi, xlab="test", ylab="BMI", main="test vs bmi") 
```


