---
title: 'STAT 500: HW9'
author: "Jasmine Mou"
date: "12/05/2017"
output: pdf_document
---

1. Using the `seatpos` data, perform a PCR analysis with `hipcenter` as the response and `HtShoes`, `Ht`, `Seated`, `Arm`, `Thigh` and `Leg` as predictors. Select an appropriate number of components and give an interpretation to those you choose. Add `Age` and `Weight` as predictors and repeat the analysis. Use both models to predict the response for predictors taking these values:
```{r echo=FALSE, results='asis', warning=FALSE}
library(knitr)
df = data.frame(matrix(c('64.800 ', '263.700 ', '181.080 ', '178.560 ', '91.440 ', '35.640 ', '40.950 ', '38.790'), nrow=1))
colnames(df)=c("Age", "Weight", "HtShoes", "Ht", "Seated", "Arm", "Thigh", "Leg")
kable(df)
```

```{r, tidy=TRUE, message=FALSE, warning=FALSE, fig.height=3}
library(pls)
rmse <- function(x,y) sqrt(mean((x-y)^2))
set.seed(123)

data(seatpos, package="faraway")
n <- nrow(seatpos)
test_seq <- sample(n, n*0.25)
grp_predictors_0 =  c("HtShoes", "Ht", "Seated", "Arm", "Thigh", "Leg")
grp_all_0 = c("hipcenter", grp_predictors_0)
grp_predictors_1 =  c("HtShoes", "Ht", "Seated", "Arm", "Thigh", "Leg", "Age", "Weight")
grp_all_1 = c("hipcenter", grp_predictors_1)

df_test = data.frame(matrix(c(181.080, 178.560, 91.440, 35.640, 40.950, 38.790, 64.800, 263.700), nrow=1))
colnames(df_test)=c("HtShoes", "Ht", "Seated", "Arm", "Thigh", "Leg", "Age", "Weight")

cal_pcr<-function(grp_all, grp_predictors){
  # assign train and test randomly for cross validation
	sp <- seatpos[, grp_all]
	p <- ncol(sp) - 1
	train_sp <- sp[-test_seq,]
	test_sp <- sp[test_seq,]

	## use pcr + RMSEP plot
	pcr_sp <- pcr(hipcenter ~., data=train_sp, validation="CV", ncomp=p)
	pcrCV_sp <- RMSEP(pcr_sp, estimate="CV")
	plot(pcrCV_sp, xlab="# of PCs", ylab="RMSEP", main=paste("# of Predictors:", p))
	pcr_sp$nc_sp <- which.min(pcrCV_sp$val)-1   
	pcr_sp$ypred_sp <- predict(pcr_sp, test_sp[,grp_predictors], ncomp=pcr_sp$nc_sp)
	pcr_sp$rmse <- rmse(pcr_sp$ypred_sp, test_sp$hipcenter)

	pcr_sp$exp <- explvar(pcr_sp)
	# pcr_sp$ld <- loadings(pcr_sp)
	# loadingplot(pcr_sp, comps=1:3, legendpos="topright")
	pcr_sp$acc <- 0
	for(i in 1:pcr_sp$nc_sp){
	  pcr_sp$acc = pcr_sp$acc + pcr_sp$exp[i]
	}
	
	return(pcr_sp)
}
par(mfrow=c(1,2))
pcr_sp_0 <- cal_pcr(grp_all_0, grp_predictors_0) # rmse = 37.34921, seed = 123
pcr_sp_1 <- cal_pcr(grp_all_1, grp_predictors_1) # rmse = 35.81502, seed = 123
pcr_sp_fit <- predict(pcr_sp_1, df_test, ncomp=pcr_sp_1$nc_sp)
```


*Perform PCR without scaling and use cross-validation over the train data to choose the number of components. Draw a RMSE vs number of components plot for visualization. *

*With only 6 predictors, we choose the number of components to be `r pcr_sp_0$nc_sp` according to the plot. The proportion of variability explained by the first `r pcr_sp_0$nc_sp` components is `r round(pcr_sp_0$acc,2)`%. The RMSE of the model over the test data is `r round(pcr_sp_0$rmse,2)`. *

*With 8 predictors, we choose the number of components to be `r pcr_sp_1$nc_sp` according to the plot. The proportion of variability explained by the first `r pcr_sp_1$nc_sp` components is `r round(pcr_sp_1$acc,2)`%. The RMSE of the model over the test data is `r round(pcr_sp_1$rmse,2)`. *

*The predicted response of the sample is `r round(pcr_sp_fit,4)`.*

<!-- *Use the prcomp() method in R to do the PCR analysis.* -->
<!-- ```{r} -->
<!-- cal_rmse_prcomp<-function(grp_all, grp_predictors, n_comp_sp){ -->
<!--   # assign train and test randomly for cross validation -->
<!-- 	sp <- seatpos[, grp_all] -->
<!-- 	train_sp <- sp[-test_seq,] -->
<!-- 	test_sp <- sp[test_seq, ] -->

<!-- 	pcr_sp <- prcomp(seatpos[-test_seq, grp_predictors], scale=TRUE) -->
<!-- 	summary(pcr_sp) -->
<!-- 	par(mfrow=c(1,2)) -->
<!-- 	biplot(pcr_sp) -->
<!-- 	plot(pcr_sp$sdev,type="l",ylab="SD of PC", xlab="PC number") -->

<!-- 	lm_pcr_sp <- lm(hipcenter ~ pcr_sp$x[,1:n_comp_sp], data=train_sp) -->

<!-- 	get_z <- function(pcr, i, test){ -->
<!-- 	  return(pcr$rotation[,i] %*% t(test[, grp_predictors])) -->
<!-- 	} -->
<!-- 	z = NULL -->
<!-- 	for (i in 1:n_comp_sp){ -->
<!-- 	  z_i = get_z(pcr_sp, i, test_sp) -->
<!-- 	  z = rbind(z, z_i) -->
<!-- 	} -->

<!-- 	ypred_sp <- coef(lm_pcr_sp) %*% rbind(1, z) -->
<!-- 	# or: ypred_sp_0 <- predict(pcr_sp, test_sp) %*% rbind(1, z) -->
<!-- 	result <- rmse(ypred_sp, test_sp$hipcenter)  -->

<!-- 	return(result) -->
<!-- } -->
<!-- ``` -->
<!-- *With first few predictors. By drawing the screeplot, we can see there is an elbow at the n=2, which is chosen as the number of components.* -->
<!-- ```{r} -->
<!-- rmse_prcomp_sp_0 <- cal_rmse_prcomp(grp_all_0, grp_predictors_0, 2) # seed=123, n=2, 166.4093, 5841.319 -->
<!-- ``` -->
<!-- *Add 2 more predictors. By drawing the screeplot, we can see there is an elbow at the n=6, which is chosen as the number of components.* -->
<!-- ```{r} -->
<!-- rmse_prcomp_sp_1 <- cal_rmse_prcomp(grp_all_1, grp_predictors_1, 6) # seed=123, n=6, 4853.191 -->
<!-- ``` -->


2. Take the `fat` data, and use the percentage of body fat, `siri`, as the response and the other variables, except `brozek` and `density` as potential predictors. Remove every tenth observation from the data for use as a test sample. Use the remaining data as a training sample building the following models:
```{r, tidy=TRUE}
data(fat, package="faraway")
n <- nrow(fat)
p <- ncol(fat) - 3
test_seq <- seq(10, n, by=10)
grp_predictors_fat = colnames(fat)[!colnames(fat) %in% c("brozek", "density")]
train_fat <- fat[-test_seq, grp_predictors_fat]
test_fat <- fat[test_seq, grp_predictors_fat]
```
(a) Linear regression with all predictors. 
```{r, tidy=TRUE}
lm_fat <- lm(siri~., data=train_fat)
rmse_lm_fat_train <- rmse(lm_fat$fit, train_fat$siri)
rmse_lm_fat_test <- rmse(predict(lm_fat, test_fat[,-1]), test_fat$siri)
```
(b) Linear regression with variables selected using AIC.
```{r, tidy=TRUE}
lmAIC_fat <- step(lm_fat)
rmse_lmAIC_fat_train <- rmse(lmAIC_fat$fit, train_fat$siri)
rmse_lmAIC_fat_test <- rmse(predict(lmAIC_fat, test_fat[,-1]), test_fat$siri)
```
(c) Principal component regression.
```{r, tidy=TRUE, fig.height=3}
pcr_fat <- pcr(siri ~., data=train_fat, validation="CV", ncomp=p)
pcrCV_fat <- RMSEP(pcr_fat, estimate="CV")
plot(pcrCV_fat, xlab="# of PCs", ylab="RMSEP", main=paste("# of Predictors:", p))
pcr_fat$nc <- which.min(pcrCV_fat$val)-1
rmse_pcr_fat_train <- rmse(predict(pcr_fat, train_fat[,-1], ncomp=pcr_fat$nc), train_fat$siri)
rmse_pcr_fat_test <- rmse(predict(pcr_fat, test_fat[,-1], ncomp=pcr_fat$nc), test_fat$siri)
```     
*From the plot we see the number of components is chosen at n=`r pcr_fat$nc`.*

(e) Ridge Regression. 
```{r, tidy=TRUE, message=FALSE}
require(MASS)
rg_fat <- lm.ridge(siri~. , data=train_fat, lambda = seq(0, 5, len=2000))
# matplot(rg_fat$lambda, coef(rg_fat), type="l", xlab=expression(lambda) ,ylab=expression(hat(beta)),col=1)
rg_fat$f_lambda <- which.min(rg_fat$GCV) # 0.01127654 
rmse_rg_fat_train <- rmse(cbind(1,as.matrix(train_fat[,-1])) %*% coef(rg_fat)[rg_fat$f_lambda,], train_fat$siri)
rmse_rg_fat_test <- rmse(cbind(1,as.matrix(test_fat[,-1])) %*% coef(rg_fat)[rg_fat$f_lambda,], test_fat$siri)
```

(f) Lasso.
```{r, tidy=TRUE, message=FALSE, fig.height=3}
require(lars)
train_fat_lsx <- as.matrix(train_fat[,-1])
train_fat_lsy <- train_fat[,1]
test_fat_lsx <- as.matrix(test_fat[,-1])
ls_fat <- lars(x=train_fat_lsx, y=train_fat_lsy)
lsCV_fat <- cv.lars(x=train_fat_lsx, y=train_fat_lsy)
ls_fat$t <- lsCV_fat$index[which.min(lsCV_fat$cv)]
# ls_fat$coef <- predict(ls_fat, type="coef", s=ls_fat$t, mode="fraction")$coef # to get the coefficients
# plot(ls_fat)
rmse_ls_fat_train <- rmse(predict(ls_fat, train_fat_lsx, s=ls_fat$t, mode="fraction")$fit, train_fat$siri)
rmse_ls_fat_test <- rmse(predict(ls_fat, test_fat_lsx, s=ls_fat$t, mode="fraction")$fit, test_fat$siri)
```

Use the models you find to predict the response in the test sample. Make a report on the performances of the models.    
*Model performance: in terms of the test data, the performance rank is PCR > LASSO > AIC > Ridge > LR.*     
```{r, tidy=TRUE}
df_report = data.frame(LR=c(rmse_lm_fat_train, rmse_lm_fat_test), AIC=c(rmse_lmAIC_fat_train, rmse_lmAIC_fat_test), PCR=c(rmse_pcr_fat_train, rmse_pcr_fat_test), Ridge=c(rmse_rg_fat_train, rmse_rg_fat_test), LASSO=c(rmse_ls_fat_train, rmse_ls_fat_test))
rownames(df_report) = c("train", "test")
kable(df_report, digits=4)
```

